Site location study for a high-mountain millimeter observatory in the Colombian Andes I: In situ data

### 2017-03-03 pasp/paperclimate.tex
some mods

Astronomical observations in the millimeter and sub-millimeter wavelength range require that atmospheric effects affecting absorption at these wavelengths are kept to a minimum \cite{chajpwv,radford2016}. The main factor contributing to the atmospheric opacity is water vapor, which very efficiently absorbs light in the THz range due to BLAHBLAHBLAH. In order to characterize a site according to its atmospheric transparency to THz radiation, it is necessary to retrieve the precipitable water vapor profile either via remote sensing techniques such as microwave radiometry \cite{radiometer}, satellite measurements, or indirectly via models of in situ climatological measurements or GPS-delay studies. \\

@@

>>>>

Besides astronomical observations, satellite communications in the THz frequency range are a growing field, given that the ``classical radio or whatevs''

Astronomical observations in the millimeter and sub-millimeter wavelength range require that atmospheric effects affecting absorption at these wavelengths are kept to a minimum \cite{chajpwv,radford2016}. The main factor contributing to the atmospheric opacity is water vapor, which very efficiently absorbs light in the THz range due to BLAHBLAHBLAH. In order to characterize a site according to its atmospheric transparency to THz radiation, it is necessary to retrieve the precipitable water vapor profile either via remote sensing techniques such as microwave radiometry \cite{radiometer}, satellite measurements \cite{spaceradio}, or indirectly via models of in situ climatological measurements or GPS-delay studies. \\

@@

 %c y después ok
